---
title: "2-PermanovaNullModels"
author: "Arlie McCarthy"
date: "25/06/2021"
output: html_document
---

ONLY RUN THIS SCRIPT IN THE SAME SESSION AS RUNNING ALL OF SCRIPT 1. IT RELIES ON PACKAGES AND OBJECTS FROM SCRIPT 1. 

#PERMANOVA - to get statistical significance for NMDS results

(are the clusterings significant?)
```{r}
#Need to make sure have community matrix and site data with matching numbers of rows
community_matrix_sites <- community_matrix %>% 
  rownames_to_column(var = "id") %>% 
  left_join(site_info)
#Re-extract community matrix and site info data now they are aligned
community_matrix_AB <- select(community_matrix_sites, -id)
community_site_info <- select(community_matrix_sites, id, area, stage, quadrat_number)
#Make a presence-absence version of community data
community_matrix_PA <- decostand(community_matrix_AB, method = "pa")
```

# Perform PERMANOVA - right now PERMANOVAs use bray-curtis distances, as the NMDS also uses the bray-curtis distance, these results are directly comparable to 
the ordination visualisation

## PERMANOVA for abundance data
```{r}
ABpermanova <- adonis2(community_matrix_AB ~ stage, community_site_info, permutations=1000,
                       method="bray")
ABpermanova
```

Cannot reject null hypothesis

So, we can detect overall differences between stages on the abundance matrix.


## PERMANOVA for presence-absence data
```{r}
PApermanova <- adonis2(community_matrix_PA ~ stage, community_site_info, permutations=1000,
                       method="bray")
PApermanova
```
CAN reject null hypothesis.

So basically: heaps of variation in abundances WITHIN each stage (cannot detect overall differences between stages), BUT just looking at the presence of species, the four stages ARE
significantly different i.e. different communities of organisms are in the four stages. 
Suggests multiple things going on - there are broad differences between the four stages in 
terms of what species are there, but there is also an area-specific effect leading to very 
variable abundances within each stage.



#Analysis of nestedness

Look at nestedness of presence-absence data:
```{r}
#Use least-constrained null model (only total count maintained)
oecosimu(community_matrix_PA, nestednodf, "r00", nsimul=1000)
```

```{r}
#Use null model where individual site richness is constrained
oecosimu(community_matrix_PA, nestednodf, "r0", nsimul=1000)
```

```{r}
#Use null model where species prevalences are constrained
oecosimu(community_matrix_PA, nestednodf, "c0", nsimul=1000)
```

```{r}
#Use null model where both site richness and species prevalences are constrained 
oecosimu(community_matrix_PA, nestednodf, "quasiswap", nsimul=1000)
```


So basically: we observe that species-poor sites tend to be nested subsets of species-rich
ones (r00 model). Allowing for the fact that some sites will be more favourable than others (r0),
we still see this pattern. Allowing for the fact that some species are inherently more common than
others (c0), we still see this pattern. However, allowing for both these things (quasiswap) we do 
NOT see this pattern: nestedness is a combined product of the suitability of certain sites and the
differential prevalence of species.