---
title: "1-DataAnalysis"
author: "Arlie McCarthy"
date: "02/05/2021"
output: html_document
---
# Code Setup and Data Import - because that is super-important 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(tidyverse)
library(readr)
library(here)
library(janitor)
```

I have one data file with the count data from lab processing. The file needs some wrangling to calculate a density (counts per cm2) for each quadrat sampled.

```{r calculating density}
taxon_count_data <- read_csv(here("data", "taxon_count_data_internal_pipework.csv"))
taxon_density <- taxon_count_data %>% 
  mutate(id = paste(area, quadrat_number, sep = "_")) %>% #some quadrats had multiple 'sample' so changed to just area and quadrat
  group_by(id, taxon, taxon_grouping_qualifier, area, stage, quadrat_number) %>% 
  mutate(density = count*multiply_by/divide_by/area_sampled*100) %>% #counts per m2
  summarise(avg_density = mean(density))
site_info <- taxon_density %>% 
  ungroup() %>% 
  select(id, area, stage, quadrat_number) %>% 
  group_by(id) %>% 
  slice(1)
```

# Preparing Data for ordination

I take my data from counts to a community matrix, with species as variables and rows as samples

```{r nMDS preparation}
community_matrix <- taxon_density %>% 
  group_by(id) %>% 
  mutate(full_taxon_group = paste(taxon,taxon_grouping_qualifier, sep = "_")) %>% 
  select(id, full_taxon_group, avg_density) %>% 
  pivot_wider(names_from = full_taxon_group, values_from = avg_density) %>% 
  ungroup() 
community_matrix[is.na(community_matrix)] <- 0
community_matrix <- community_matrix %>% 
  select(-"mussel_valves intact, no tissue", -"NA_NA", -"mussel_valves intact") %>% 
  clean_names() %>% 
  group_by(id) %>% 
  mutate(mussel = sum(mussel_na +
                      mussel_0_5cm +
                      mussel_2_5mm +
                      mussel_5mm +
                      mussel_large +
                      mussel_0_3_3cm_with_tissue +
                      mussel_0_3cm_with_tissue +
                      mussel_0_5cm_2 +
                      mussel_intact_shells_2cm +
                      mussel_with_tissue_0_5cm),
         acorn_barnacle = sum(acorn_barnacle_na +
                              acorn_barnacle_intact),
         clam = clam_na +
           clam_0_2cm,
         c_f_jassa = c_f_jassa_na +
           c_f_jassa_heads +
           c_f_jassa_large +
           c_f_jassa_medium +
           c_f_jassa_small +
           green_eye_amphipod_na,
         acorn_barnacle_ribbed = acorn_barnacle_ribbed_na +
           barnacle_type_2_na,
         sponge = sponge_na +
           tiny_sponges_na,
         crab = crab_na +
           juvenile_crab_na,
         encrusting_bryozoan_fragment = encrusting_bryozoan_fragment +
           encrusting_bryozoan_na,
         white_tufted_bryozoan_fragment = white_tufted_bryozoan_fragment +
           white_tufted_bryozoan_na,
         other_bivalve = other_bivalve_na +
           other_bivalves_na,
         white_tufted_bryozoan_colony = white_tufted_bryozoan_colony +
           white_tufted_bryozoan_large_colony +
           white_tufted_bryozoan_small_colony,
         polychaete_tube_fragment = polychaete_tube_fragment +
           polychaete_tube_na,
         gastropod = gastropod_na +
           gastropods_na,
         .keep = "unused") %>% 
  ungroup() %>% 
  column_to_rownames(var = "id") %>% 
  filter_all(any_vars(. != 0))
```

## nMDS

```{r nMDS}
internal_pipe_nmds <- metaMDS(community_matrix, k = 2, trymax = 500)
internal_pipe_nmds
data_scores <- as.data.frame(scores(internal_pipe_nmds))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data_scores$site <- rownames(internal_pipe_nmds)  # create a column of site names, from the rownames of data.scores
data_scores <- data_scores %>% 
  rownames_to_column(var = "id") %>% 
  left_join(site_info)
data_scores$n_species <- specnumber(community_matrix, MARGIN = 1)
data_scores$shannon <- diversity(community_matrix, index = "shannon")
data_scores$simpson <- diversity(community_matrix, index = "simpson")
species_scores <- as.data.frame(scores(internal_pipe_nmds, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species_scores$species <- rownames(species_scores)

```

Plotting with ggplot
```{r}
ggplot() + 
  #geom_text(data=species_scores,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(data=data_scores,aes(x=NMDS1,y=NMDS2,shape=as_factor(stage),colour=n_species),size=3) + # add the point markers
  #geom_text(data=data_scores,aes(x=NMDS1,y=NMDS2,label=area),size=6,vjust=0) +  # add the site labels
  coord_equal() +
  theme_bw()
```
Checking the nMDS

```{r}
goodness(internal_pipe_nmds) # Produces a results of test statistics for goodness of fit for each point

stressplot(internal_pipe_nmds) # Produces a Shepards diagram
```